---
title: "risk of CS among multi-parous women"
author: "Eskedar and Nigus"
date: "2023-12-09"
output: html_document
---


##Load important pakage 
```{r}
library(PredictABEL)
library(pROC)
library(readxl)
library(haven)
library(boot)
library(ggplot2)
library(tidyverse)
library(MASS)
```

## import the data
```{r}
library(haven)
dat_cs <- read_sav("thesis_data_NB.sav")
```

## explore the data
```{r}
str(dat_cs)
dat_cs <- dat_cs[-310, ]
```

## filter the data
```{r}
new_data <- dat_cs %>% 
  dplyr::select(participant_id, age, gravidity, parity, abortion, IUFD, Preterm_d, instrumental_d, C_S, ANC, mode_delivery_last, DM, PROM, PTL, APH, HDP, mode_delivery,medical_illness,number_ANC_cat, residence)
```

## explore the new data
```{r}
colSums(is.na(new_data))
str(new_data)
```

## recode variables
```{r}
## outcome
new_data <- mutate(new_data, outcome= ifelse(as.character(new_data$mode_delivery) == "5", 1, 0))
##Advanced age
new_data<-mutate(new_data, Advanced_age=ifelse(age>34,1,0))
##history of any medical illness 
new_data$medical_illness<-ifelse(as.character(new_data$medical_illness)=="1", 1, 0)
##ANC visit
new_data$number_ANC_cat<-ifelse(as.character(new_data$number_ANC_cat)=="1", 1, 0)
## Abortion
new_data$abortion <-ifelse(as.character(new_data$abortion) == "1", 1, 0)

## IUFD
new_data$IUFD <-ifelse(as.character(new_data$IUFD) == "1", 1, 0)

## PRETERM
new_data$Preterm_d <-ifelse(as.character(new_data$Preterm_d) == "1", 1, 0)

## inst delivery
new_data$instrumental_d <-ifelse(as.character(new_data$instrumental_d) == "1", 1, 0)

## CS
new_data$C_S <-ifelse(as.character(new_data$C_S) == "1", 1, 0)

## ANC
new_data$ANC <-ifelse(as.character(new_data$ANC) == "2", 1, 0)

## previous mode del
new_data$mode_delivery_last <-ifelse(as.character(new_data$mode_delivery_last) == "2", 1, 0)

## PROM
new_data$PROM <-ifelse(as.character(new_data$PROM) == "1", 1, 0)

## HDP
new_data$HDP <-ifelse(as.character(new_data$HDP) == "1", 1, 0)

## Parity
new_data <- mutate(new_data, parity_cat= ifelse(parity > 4, 1,0))

#APH
new_data$APH <- ifelse(as.character(new_data$APH) == "1", 1,0)
```
 
 ## Descriptive table
```{r}
table1(~ as.factor(Advanced_age) + as.factor(medical_illness) + as.factor(number_ANC_cat) + as.factor(abortion) + as.factor(IUFD) + as.factor(Preterm_d) + as.factor(C_S) + as.factor(ANC) + as.factor(mode_delivery_last) + as.factor(PROM) + as.factor(HDP) + as.factor(parity_cat) + as.factor(APH) | as.factor(outcome), data= new_data) 
```
 
 
 ## Convert variables to factor
```{r}
new_data %>% mutate_at(c('Advanced_age', 'HDP', 'PROM', 'Preterm_d', 'medical_illness', 'APH', 'abortion', 'number_ANC_cat', 'IUFD', 'ANC', 'mode_delivery_last', 'parity_cat', 'APH', 'outcome'), as.factor)

```
 
 
 ## Table by outcome characteristics
```{r}

new_data1 <- new_data %>% select(participant_id, Advanced_age, HDP, PROM, Preterm_d, medical_illness, APH, abortion, number_ANC_cat, IUFD, ANC, mode_delivery_last, parity_cat, APH, outcome)

Table_1 <- new_data1 %>% 
  dplyr::select(- participant_id) %>% 
  tbl_summary(by= outcome) %>% 
  add_overall() 

print(Table_1)
```
 
 
```{r}
full_fit <- glm(outcome~ Advanced_age + abortion + parity_cat + mode_delivery_last + number_ANC_cat + medical_illness + HDP + APH, family= binomial, data = new_data)

summary(full_fit)

## OR for full model
OR_full <- exp(cbind(OR= coef(full_fit), confint(full_fit)))
round(OR_full, digits = 3)

## model reduction

stepAIC(full_fit)
```

## reduced model
```{r}
red_model <- full_fit <- glm(outcome~  HDP + APH + mode_delivery_last, family= binomial, data = new_data)

## ORs and CIs
OR_CI <- exp(cbind(OR= coef(red_model), confint(red_model)))
round(OR_CI, digits = 3)
```

## model performance/calibration 
```{r}
predicted_probabilities <- predict(red_model, type = "response")
calibration_data <- data.frame(Outcome = new_data$outcome, Predicted_Risks = predicted_probabilities)
predRisk <- predRisk(red_model)
plotCalibration(calibration_data, 1, predRisk, 10,plottitle = "Calibration_of_the_model")
```
## Model performance/descrimination plot
```{r}
predsred = predict(red_model,type="response")
plot.roc(new_data$outcome, predsred,print.auc=TRUE)
```
##Model development using Least square absolute shrinkage operator 
```{r}
## LASSO regression

# Extract response variable 'outcome'
y <- new_data$outcome

# Select predictor variables
x <- model.matrix(outcome ~ parity + abortion + IUFD + PROM + HDP + mode_delivery_last + ANC, data = new_data)[,-1]

set.seed(123)  # for reproducibility
cv_fit <- cv.glmnet(x, y, family = "binomial")

print(cv_fit$lambda.min)

# Predict using the fitted model
predictions <- predict(cv_fit, newx = x, s = "lambda.min", type = "response")

# Get coefficients at the lambda that gives minimum mean cross-validated error
lasso_coefficients <- coef(cv_fit, s = "lambda.min")

# Print the coefficients
print(lasso_coefficients)
```

## Internal validation using bootstraping 
```{r}
# load packages
library(boot)
library(pROC)

# Step 2: Define your prognostic model
## Assuming you want to build a logistic regression model
model <- glm(outcome ~ HDP + APH + mode_delivery_last, family = binomial, data = new_data)

# Step 3: Define a function to compute the AUC
auc <- function(new_data, indices) {
  d <- new_data[indices, ]
  model_fit <- glm(outcome ~ HDP + mode_delivery_last, family = binomial, data = d)
  predicted <- predict(model_fit, newdata = d, type = "response")
  observed <- d$outcome
  return(pROC::auc(pROC::roc(observed, predicted)))
}
# Step 4: Perform bootstrap resampling
boot_results <- boot(data = new_data, statistic = auc, R = 1000)

# Step 5: Extract the AUC values for each bootstrap sample
boot_auc <- boot_results$t
print(boot_auc)
```

## Calculate overoptimsm coefficient 
```{r}
apparent_performance <- 0.777 
true_perf <- mean(boot_auc)  # mean of the bootstrap sample
over_opt_coef <- (apparent_performance - true_perf) / apparent_performance  # 0.04 --> low chance of overoptimism 
```


## Risk_score calculation 
```{r}
# convert to numeric
new_data$HDP_rs<-as.numeric(new_data$HDP)
new_data$mode_delivery_last_rs<-as.numeric(new_data$mode_delivery_last)
new_data$APH_rs <- as.numeric(new_data$APH)

# calculate total risk score
new_data$Risk_score<-(1*new_data$HDP_rs)+(3*new_data$mode_delivery_last_rs) + (2*new_data$APH_rs) 
cor.test(new_data$outcome,new_data$Risk_score)
```

## # calculate probability for corresponding risks
```{r}
length(new_data$participant_id[new_data$Risk_score=="0" & new_data$outcome=="1"])/ length(new_data$participant_id[new_data$Risk_score=="0"]) # risk 0.06
length(new_data$participant_id[new_data$Risk_score=="1" & new_data$outcome=="1"])/ length(new_data$participant_id[new_data$Risk_score=="1"]) # risk 0.15
length(new_data$participant_id[new_data$Risk_score=="2" & new_data$outcome=="1"])/ length(new_data$participant_id[new_data$Risk_score=="2"]) # risk 0.38
length(new_data$participant_id[new_data$Risk_score=="3" & new_data$outcome=="1"])/ length(new_data$participant_id[new_data$Risk_score=="3"]) # risk 0.6
length(new_data$participant_id[new_data$Risk_score=="4" & new_data$outcome=="1"])/ length(new_data$participant_id[new_data$Risk_score=="4"]) # risk 1
length(new_data$participant_id[new_data$Risk_score=="5" & new_data$outcome=="1"])/ length(new_data$participant_id[new_data$Risk_score=="5"]) # risk 0.5


```


##model with the risk score 
```{r}
risk<-glm(new_data$outcome~new_data$Risk_score,family = binomial)
predsrisk = predict(risk,type="response")
plot.roc(new_data$outcome, predsrisk,print.auc=TRUE)
```
```{r}
library(dcurves)
library(rmda)
Dcred<-decision_curve(outcome~  HDP + mode_delivery_last, family= binomial, data = new_data,thresholds = seq(0, 1, by = .01),confidence.intervals = 0.95) 
# Define the colors for each line
line.colors <- c("blue", "green","red")
plot_decision_curve(x=Dcred, curve.names=c("Model", "Intervention for all", "Intervention for none"), cost.benefit.axis = TRUE,col = line.colors)
```

